Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-05 paper/Intuition.tex
intuition

\begin{inv}
At any time $t$, a cache $c$ which is clean for an address $a$ will have the most
up-to-date value for that address, \ie{}
\begin{enumerate}
\item $c.\data[a] = \initData(a)$ and no store request $q'$ has been processed at
any time $t'$ such that $t' < t \wedge a = q'.\addrQ$, or
\item There is a store request $q$ that was processed at time $t_q$ such that
$t_q < t \wedge a = q.\addrQ \wedge c.\data[a] = q.\dataQ$ and no other store
request $q'$ was processed at any time $t'$ such that $t_q < t' < t \wedge a =
q'.\addrQ$.
\end{enumerate}
\label{latestValue}
\end{inv}
A clean cache is defined as follows.
\begin{defn}
\textit{clean}: A cache is said to be \clean{} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for all its children is at most \Sh.
\end{defn}

@@

Once a response has been produced for a request $i$, the cache increments the
counter for the request it processes next. Thus there is at most one response
to each request, satisfying Definition~\fullref{uniqRespLabels}. As leaf caches
have no children, a leaf cache in at least \Sh{} state for address $a$ is
\clean{} for that address. Since processing a load request requires the cache
to be at least in the \Sh{} state, its response, which supplies the data for
that address present in the cache, supplies the latest value by
Invariant~\fullref{lastestValue}.  Thus, the final condition for
Theorem~\fullref{storeAtomicity} is also satisfied. All that remains is to show
how these invariants hold.

@@

Correctness of Invariants \fullref{uniqM} and \fullref{latestValue} can be
proved using the following invariants on a parent-child pair.

@@

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair, for the same address.
\end{inv}

>>>>

\begin{inv}
\textit{latestValue}:
At any time $t$, a cache $c$ which is clean for an address $a$ will have the most
up-to-date value for that address, \ie{}
\begin{enumerate}
\item $c.\data[a] = \initData(a)$ and no store request $q'$ has been processed at
any time $t'$ such that $t' < t \wedge a = q'.\addrQ$, or
\item There is a store request $q$ that was processed at time $t_q$ such that
$t_q < t \wedge a = q.\addrQ \wedge c.\data[a] = q.\dataQ$ and no other store
request $q'$ was processed at any time $t'$ such that $t_q < t' < t \wedge a =
q'.\addrQ$.
\end{enumerate}
\label{latestValue}
\end{inv}
A clean cache is defined as follows.
\begin{defn}
\textit{clean}: A cache is said to be \clean{} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for all its children is at most \Sh.
\end{defn}

@@

Once a response has been produced for a request $i$, the cache increments the
counter for the request it processes next. Thus there is at most one response
to each request, satisfying Definition~\fullref{uniqRespLabels}. As leaf caches
have no children, a leaf cache in at least \Sh{} state for address $a$ is
\clean{} for that address. Since processing a load request requires the cache
to be at least in the \Sh{} state, its response, which supplies the data for
that address present in the cache, supplies the latest value by
Invariant~\fullref{latestValue}.  Thus, the final condition for
Theorem~\fullref{storeAtomicity} is also satisfied. All that remains is to show
how these invariants hold.

@@

Correctness of Invariants \fullref{localCompatibility} and \fullref{latestValue} can be
proved using the following invariants on a parent-child pair.

@@

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair, for the same address.
\label{respFIFO}
\end{inv}
